\ttitle

### 2013-02-04 Chapters/Chapter1.tex
Added gitignore file

>>>>

*** Most of the activity recognition techniques can be classified in two 
different types. Some rely on a state-space model, in which the activities to 
be recognized are represented in a statistical manner. Bayesian networks, 
Finite State Machines and Hidden Markov Models can be considered among those. 
Other techniques process the directly as a pattern recognition task, such as 
Support Vector Machines, Neural Networks, Dynamic Time Warping, and Bayes and 
K-means clustering.

### 2013-02-04 Chapters/Chapter2.tex
Added gitignore file

\subsection{Hidden Markov Models}
Viterbi algorithm / path; find the most likely sequence of cause-states for 
the observed measures.

>>>>

\subsection{Hidden Markov Models}
Many dynamic systems can be viewed as some sequence of consecutive states 
which are observable by the world. These kind of systems are known as Markov 
Chains. When the state in which the system is can not be observed with 
complete certainty, a system known as a Hidden Markov Model (HMM) can be 
constructed. It has been shown that HMMs are effective on in the task of 
speech recognition \cite{rabiner1989tutorial}. Because of some resemblance 
between natural speech construction and human activities, modifications have 
been able to classify human activities \cite{guenterberg2009distributed} and 
extensions are implemented which create layers of HMMs to handle different 
granularity over time and reduce retraining 
\cite{oliver2002layered,perdikis2008recognition}. A HMM is a state-space model 
representing a single model for each label which needs to be classified.

Formally, a Hidden Markov Model is characterized by the following elements 
\cite{rabiner1989tutorial}:
\begin{enumerate}
	\item A number of $N$ states in the model. These states are a 
	representation of the underlying mechanism of the system and can not be 
	directly observed from outside (hence the \emph{hidden} model). Individual 
	states are referenced by $S = \{S_1, S_2, \dots, S_N \}$ and a single 
	state in time $t$ is denoted by $q_t$.
	\item An alphabet of $M$ discrete observable symbols for each state. An 
	observed symbol is the only indication of the systems current situation. 
	Each individual symbol is referenced by $V = \{ v_1, v_2, \dots, v_m \}$.
	\item A state transition probability distribution $A = \{ a_{ij} \}$ 
	indication the probability of being in state $j$ at time $t+1$ when the 
	state on time $t$ was $i$:
	\begin{eqnarray}
		a_{ij} = P [ q_{t+1} = S_j | q_t = S_i ], & 1 \le i, j \le N
	\end{eqnarray}
	Note that by setting the probability $a_{ij}=0$ for some $i$ and $j$ 
	indicates that $j$ is not reachable from state $i$.
	\item A symbol observation probability distribution in state $j$, $B = \{ 
	b_j(k) \}$, where
	\begin{eqnarray}
	b_j(k) = P[ v_k\ \mathrm{at} \ t | q_t = S_j], & 1 \le j \le N \nonumber \\
	& 1 \le k \le M
	\end{eqnarray}
	Note that every symbol can be observed in every state.
	\item The initial state distribution in which the system starts $\pi= \{ 
	\pi_i \}$ where
	\begin{eqnarray}
	\pi_i = P[q_1 = S_i], & 1 \le i \le N
	\end{eqnarray}
\end{enumerate}
This system is often abbreviated to
\begin{equation}
	\lambda = \{ \pi_i, a_{ij}, b_j(k)\}
\end{equation}
The HMM will act as a generator giving a sequence of $T$ observations, each 
being a symbol from $V$:
\begin{equation}
O = O_1, O_2 \dots O_T
\end{equation}

Given the characteristics of a specific HMM, there are three major application 
which serve the classification of an activity. The first application is, given 
a sequence of observations $O = O_1,O_2,\dots,O_T$, to label this sequence 
with the activity with 
the highest associated probability for that label, e.g. compute $P(O|\lambda)$ 
for each model $\lambda$. The second task a HMM can perform is to 
\emph{explain} a observation sequence $O$, by choosing an optimal 
corresponding state sequence $Q = q_1,q_2,\dots,q_T$. The last application is 
more of a problem statement, being the task to train the HMM to adjust the 
parameters $\lambda = \{ \pi_i, a_{ij}, b_j(k)\}$ to maximize $P(O|\lambda)$. 
In the overall task of activity recognition, one would first use the last 
application to train a set of HMMs with the provided and labeled training 
data. The second application can be used to further optimize the HMM (set the 
number of states, adjust vector codebook etc.) and improve the capability of 
modeling the activities. Finally, when the system is provided with unlabeled 
data the first application is used to identify the observed sequence with 
trained examples, by comparing the probability result of multiple HMMs.

Viterbi algorithm / path; find the most likely sequence of cause-states for 
the observed measures.
